How to Learn Useful Changes to Software Projects\\(to Reduce Runtimes and Software Defects)

### 2015-08-24 Manuscript.tex
Updates from ShareLaTeX

This presents such a tool, called XTREES and evaluates it against three alternatives.
In our experiments, XTREES reduced
 the expected values of defect counts and  runtimes of software projects to    
 45\%,61.5\%  (median) and 22\%,6\% (best case) of the initial values, respectively.
\end{abstract}
\section{Introduction}
Business users are demanding  tools that support    business-level interpretations of their data. At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}. Panelists commented  ``prediction is all well and good, but what about decision making?''. Note that these panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself. So:
\bi
\item
Instead of just accepting {\em predictions} on how many software defects to expect, business users might now demand a {\em plan} to reduce the likelihood of those defects.
\item
Instead of just accepting {\em predictions} on the runtime time of their software, business users might now demand a {\em plan} to reduce that runtime. 
\ei
In response to this business-level demands for planners,
we propose a novel {\em planning} approach called XTREES for learning changes to a software system
such that its performance ``improves'', according to some measure. This paper uses XTREES
to reduce  the expected value of:
\bi
\item  Defects     in Jureczko et al.'s    JAVA systems~\cite{jureczko10}.
\item Runtimes   in   software    configured by  Siegmund et al.~\cite{sven12}.
\ei 
The approach discussed here is novel since most data miners used in software analytics
comment on ``what is'' and not ``what to do''. For example, a clustering
algorithm might find clumps of similar examples, but does not comment on the minimal
change(s) required to move from clump to clump.

The contributions of this paper are two-fold:
\be
\item
The  XTREES planner, which has not been previously published. For demonstration purposes, we show XTREES
working to reduce software runtimes and software defects. That said, the method is quite general to any table of data with weighted class values (so that XTREES
can determine that some rows are preferred and some are to be avoided).  
\item An evaluation strategy used to comparatively assess planners. As  shown below, that strategy less us identify  inferior planners. With this strategy, other researchers could assess improvements or alternatives to the XTREE planner. 
\ee
In all, this paper explores and comparatively assesses four methods for learning planners (and XTREES is Method\#4).
As an historical note,  Method\#2 was previously proposed in one of our prior publications~\cite{me12c}. As shown below,
that method is a comparatively far less effective than XTREES. Hence, we   recant that prior work and, based on the results of this
paper, we now endorse  XTREES.

The rest of this paper is structured as follows.
After describing our test data,
this paper  walks the reader through four methods that start
with a standard clustering algorithm, the  progresses on to XTREES. The second
half of the paper then presents an extensive evaluation of these methods using 
the Jureczko  and Siegmund data sets. This is followed by notes on related work
and validity.

@@

\subsection{Trusting the Changes}\label{sect:trust}
   XTREES is evaluated by  comparing
predicted performance scores before and after a planner makes changes to the feature values of an example:
After making those
changes, we may have a new example that has never seen before. Therefore, it must be asked:
\bi
\item
{\em Can we trust the predictions made on such new examples?}
\ei
Our answer is ``yes'', if those changes select for the {\em region of competency} of the predictor. When we assess
any data miner, there are two ``clouds'' of data:
\bi
\item One cloud is the space of training examples.
\item Another cloud is the space of test examples.
\ei
For a visualization of these clouds, see \fig{howxy}.

@@

Note that predictions  about changes that  fall within the space of the training+test data, will be at least
as reliable as the performance value found in the data mining experiment.
Given this, one thing  can be asserted about predictions on changed examples:
\bi
\item Predictions about changes that move examples towards/away from the cloud of training data can be trusted more/less (respectively).
\ei  
Accordingly, all the methods of this paper make {\em trust-increasing} recommendations. That is, they push change examples {\em towards} the
training examples.  To see how this works, 
 \fig{howxy} is from the {\em ivy} data
set, which is one of the Jureczko data sets explored in this paper. It shows: (1)~the training examples, (2)~the test examples, and (3)~the
changed  examples displaced after applying one of our planning methods.
To generate that figure, we placed all of 1+2+3 into the same data set then computed the first
two  components (using PCA). Note that the  the   changed examples
cases  (shown in green)  fall closer to the training cases (shown in gray) than
the test cases (shown in red).  That is, if we believe
that the gray data is  suitable for assessing
the test cases (as indicated by the {\em ivy} results from \fig{j}), then it follows
that we believe even more in the predictions
generated from the displaced data).

This pattern of \fig{howxy} (where the changes examples are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
paper. Hence,  we can assert that
predictors learned from these training data have some authority in the regions
used to generate the plans\footnote{If the reader is concerned that this 2d projection
losses important information, then they should note that two separate  studies~\cite{papa13,divya15}
have confirmed that building defect predictors from this 2d space are just as effective
as predictors learned using all the defect attributes of \fig{ck}.}.

\begin{figure}[!t]
  \includegraphics[width=2in,bb=0 0 100 100]{figs/aa.eps} 
 % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
\caption{Training examples (in gray), test examples (in red) and 
changed examples generated by a planner (in green). Note how the green (changed) examples are closer to the  training
date than the original test data
i.e. predictors trained on the training examples and
applied to the test examples should
work just as well, or better, on the changed test examples.}\label{fig:howxy}
\end{figure}

@@

That said, the above comes with two important caveats:
\bi
\item 
We   strongly recommend that predictors are assessed prior to planning. That
issue is explored further in the next section. 
\item
Where possible, planners should be assessed via some external
oracle that can accurately assess new examples. For an example of that kind of analysis,
see the Validity section, later in this paper.
\ei

  \section{Test Data}\label{sect:tesd}
To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
  software system   configuration data from by  Siegmund et al.~\cite{sven12}.
  See github.com/ai-se/XTREES\#data for full access to this data.

@@

\section{Four Planning Methods}
In his textbook on empirical methods, Cohen~\cite{cohen95} advises that any supposedly better
algorithm should be baselined against simpler alternatives. Accordingly,
this section described XTREES (which we call Method4) against three  
alternatives: 
  \bi
  \item  
  Method1, Method2 and Method3 use the   top-down
	bi-clustering method described in \fig{where}.C  which recursively divides the
	data in two  using a dimension that captures the greatest variability in the data. 
	\item
 Method4 uses the decision tree learner of \fig{where}.D. 
  \ei
  Note that all our methods have the two properties proposed in \tion{prelim}:
  \bi
  \item They are {\em local learners}; i.e. different test examples
  will be given plans that are specialized  to their particulars;
  \item They  are {\em trust-increasing} in the manner described in \tion{trust};
  i.e. they change changing examples such that they move {\em closer} to the training data.
  \ei

@@

\subsection{Method4 =  XTREES =\\ Deltas in Decision Tree Branches}

@@

{\em Details:} One potential problem with Methods 1,2 and 3 is the {\em unsupervised} nature of
the algorithm used for the initial clustering. Unsupervised methods like WHERE
execute without knowledge of the target class.
{\em Supervised} methods, on the other hand, are allowed to reflect on more knowledge including
the values in the target class. In theory, this extra knowledge can lead to better data mining.
Accordingly, XTREES uses a supervised method to generate its plans.
Specifically, it  uses the decision tree algorithm of \fig{where}.D to divide the data.
Next, XTREES builds plans from the branches of the decision trees. It does so by asking three questions:
\be
\item
What {\em current} branch does a test case fall in?
\item What {\em desired} branch would the test case want to move to?
\item What are the {\em deltas} between current and desired? 
\ee
The last question can be answered by finding the deltas in branches of the decision tree that lead to {\em desired} from {\em current}. For full algorithmic details, refer to \fig{xtrees_bare}.

@@

\begin{figure}[!b]
\small 
\[
\begin{array}{r}
\mathrm{project}\\
\mathrm{data}
\end{array} 
\left\{\begin{array}{l}\mathit{train}
        \left\{\begin{array}{l}
                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; XTREES)}
              \end{array}\right.
       \\
      ~\\
\mathit{test}  
    \left\{\begin{array}{l@{~}l}
           \mathit{before}& =\mathrm{Performance\; scores \hspace{2pt} in \hspace{2pt}}\mathit{test}\\
           \mathrm{\bf if\;}\mathit{before} & >  \mathit{0}\\
           \mathrm{\bf then} &
           \left\{
            \begin{array}{l}
                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
                \mathrm{{\bf return}\;} R=\frac{\mathit{after}}{\mathit{before}}
            \end{array}
          \right.
   \end{array}\right.
\end{array} \right. 
\]

@@

\subsubsection{Studying Repeated Samples}
Our methods contain some stochastic algorithms; e.g. WHERE's selection of ``what example to explore first'' (see \fig{where}.C) and
  XTREES' occasional use of a random guess when deciding what part of a discretized range to include in the plan
  (see \fig{xtrees_bare}). Hence, we report the $R$ values seen in 40 repeated runs
  (with different random number seeds).
The the value 40 was chosen to be  larger than the 30 samples  required
to satisfy the central limit theorem.

@@

The above table has a  left-hand-side  {\bf Rank} column. This column reports if
the values for each method are statistically different and are more than trivially different. 
In this example, table, CD and BIC are ranked together while XTREES and CD+FS are ranked best and worst, respectively.

@@

XTREES performs well in terms of  effectiveness. 
\fig{jur} and \fig{conf} show the effectiveness of our methods seen in 40 repeats with each data set.
In these experiments,   the dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively. Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data.
\bi
\item For Siegmund data, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data. 
\item For the Jureczko data, we used the training and testing sets of \fig{j}. For these datasets,
 all the SMOTE-ing and Random Forst tunings (discussed in \tion{tesd})
occurred in the {\em train} phase of \fig{work}.
\ei
The results show planning can be very effective-- some of the expected value of the reductions are very large:
\bi
\item
Down to 22\% of the original baseline in Ant of \fig{jur};
\item
Down to 22\% of the original baseline in X264  of \fig{conf};
\ei
In all these results,  with one exception, XTREES always is the top-ranked method and with a median improvement at least 10\% lower (i.e. better) than the next ranked method. The  exception is SQL which (as shown in \fig{conf}), seems to defy all methods for optimization
so XTREES can only reduce the expected values of the runtimes by 1\%.

@@

As predicted in the above discussion on Method 1 (CD), this method always proposes changes  to most features. Method3 (BIC) often proposes changes to many features, which 
would make its plan tedious to implement. 
Method 2 (CD+FS)
proposes less, due to its use of feature selection and, in the case of {\em BDBJ}, it fixates on changing just one feature (which
probably explains why CD+FS performs so poorly for {\em BDBJ} in \fig{conf}). Finally, Method4 (XTREES) rarely fixates
on changing one feature (only on {\em LUCENE}'s lines of code feature) and seems to prefer making many small changes,
all over the space of features.

\fig{changed} reports the changes just for   {\em lucene}) and {\em BDBJ}. \fig{types} shows the mean changes for features found by our
four methods (here, mean change is the average height of the histogram bars seen in, say, \fig{changed}). Overall, Method1 (CD) and
Method3 (BIC) make the most changes while XTREES and CD+FS make the fewest.

@@

This main contribution of this paper was to  define an evaluation strategy for this kind of reasoning, and to show that (using that strategy) we
can (a)~select useful planning methods while (b)~culling clearly
inferior planning methods. 
We apply this strategy to XTREES and three alternate planning methods to find:
\bi
\item Two methods that are clearly inferior;
\item One method that proposes complex changes, which work as well as XTREES;
\item And XTREES, that proposes very simple changes.
\ei
Of course, there are many more methods for generating plans and
no   one paper can survey them all.
But the goal of this paper is not to claim that (e.g.) XTREES is some absolute optimal algorithm. Rather, it is
to offer a baseline result (with XTREES) and an  evaluation strategy that  can assess  if alternate methods are better than XTREES.
Our hope is that other   researchers apply this strategy  to repeat and/or improve
and/or refute our results. To that end, all our materials and data sets are available on-line, see
github.com/ai-se/XTREES.

@@

Using DTREE, the test cases can be categorized into one of the branches of the tree. Now, to generate the contrast sets, we determine (1) What \textit{current} branch does a test case fall in?; (2) What \textit{desired} branch would the test case have to move to?; (3) What are the deltas between \textit{current} and \textit{desired}? The last question can be answered by finding the deltas in branches of the decision tree that lead to \textit{desired} from \textit{current}. For full algorithmic details, refer to \fig{xtrees_bare}.

A tree structure such as DTREE is characterized by attributes such as size and depth. It is worth noting that these attributes have a profound impact on its performance. Our initial motivation for using DTREE was that they could serve as a medium for experts to reason about a data. The size of the tree, if too large, jeopardizes the readability of the solutions by increasing the complexity. We have, therefore, endeavored to reduce the size of the tree by pruning away irrelevant solutions that do not contribute to better solutions (refer to step-2 of \fig{xtrees_bare}).

>>>>

This presents such a tool, called XTREE and evaluates it against three alternatives.
In our experiments, XTREE reduced
 the expected values of defect counts and  runtimes of software projects to    
 45\%,61.5\%  (median) and 22\%,6\% (best case) of the initial values, respectively.
\end{abstract}
\section{Introduction}
Business users are demanding  tools that support    business-level interpretations of their data. At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}. Panelists commented  ``prediction is all well and good, but what about decision making?''. Note that these panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself. So:
\bi
\item
Instead of just accepting {\em predictions} on how many software defects to expect, business users might now demand a {\em plan} to reduce the likelihood of those defects.
\item
Instead of just accepting {\em predictions} on the runtime time of their software, business users might now demand a {\em plan} to reduce that runtime. 
\ei
In response to this business-level demands for planners,
we propose a novel {\em planning} approach called XTREE for learning changes to a software system
such that its performance ``improves'', according to some measure. This paper uses XTREE
to reduce  the expected value of the
defects     in Jureczko et al.'s    JAVA systems~\cite{jureczko10};
and the runtimes   in   software    configured by  Siegmund et al.~\cite{sven12}.

The approach discussed here is novel since most data miners used in software analytics
comment on ``what is'' and not ``what to do''. For example, a clustering
algorithm might find clumps of similar examples, but does not comment on the minimal
change(s) required to move from clump to clump.

The contributions of this paper are two-fold. Firstly, we offer
   XTREE, which has not been previously published.
   XTREE     works  on table of data with weighted class values indicating what rows are ``good'' and what rows are ``bad''.
XTREE returns plans that are changes to feature values in the ``bad'' rows to make them more ``good''.
Secondly, an evaluation strategy used to comparatively assess planners. As  shown below, that strategy less us identify  inferior planners. For example, when exploring software runtimes and software defects,
we show that XTREE is significantly better than the methods proposed in two of our prior publications~\cite{me12c}.

The rest of this paper is structured as follows.
After describing our test data,
this paper  walks the reader through four methods that start
with a standard clustering algorithm, the  progresses on to XTREE. The second
half of the paper then presents an extensive evaluation of these methods using 
the Jureczko  and Siegmund data sets. This is followed by notes on related work
and validity.

@@

\subsection{Trusting the Changes}\label{sect:trust}
   XTREE is evaluated by  comparing
predicted performance scores before and after a planner makes changes to the feature values of an example:
After making those
changes, we may have a new example that has never seen before. Therefore, it must be asked:
\bi
\item
{\em Can we trust the predictions made on such new examples?}
\ei
Our answer is ``yes'', if those changes select for the {\em region of competency} of the predictor. When we assess
any data miner, there are two ``clouds'' of data:
\bi
\item One cloud is the space of training examples.
\item Another cloud is the space of test examples.
\ei
For a visualization of these clouds, see \fig{howxy}.

\begin{figure}[!t]
  \includegraphics[width=\linewidth]{figs/twodee.eps} 
 % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
\caption{Training examples (in gray), test examples (in red) and 
changed examples generated by a planner (in green). Note how the green (changed) examples are closer to the  training
date than the original test data
i.e. predictors trained on the training examples and
applied to the test examples should
work just as well, or better, on the changed test examples.}\label{fig:howxy}
\end{figure}

@@

Note that predictions  about changes that  fall within the space of the training+test data, will be at least
as reliable as the performance value found in the data mining experiment.
Given this, one thing  can be asserted about predictions on changed examples:
\bi
\item Predictions about changes that move examples towards/away from the cloud of training data can be trusted more/less (respectively).
\ei  
Accordingly, we should use  {\em trust-increasing} planners. That is, we should use planners that push changde examples {\em towards} the
training examples.  To see how this works, 
 \fig{howxy} is from the {\em ivy} data
set, which is one of the Jureczko data sets explored in this paper. It shows: (1)~the training examples, (2)~the test examples, and (3)~the
changed  examples displaced after applying one of our planning methods.
To generate that figure, we placed all of 1+2+3 into the same data set then computed the first
two  components (using PCA). Note that the  the   changed examples
cases  (shown in green)  fall closer to the training cases (shown in gray) than
the test cases (shown in red).  That is, if we believe
that the gray data is  suitable for assessing
the test cases (as indicated by the {\em ivy} results from \fig{j}), then it follows
that we believe even more in the predictions
generated from the displaced data).

(Aside: If the reader is concerned that this 2d projection
losses important information, then they should note that two separate  studies~\cite{papa13,divya15}
have confirmed that building defect predictors from this 2d space are just as effective
as predictors learned using all the defect attributes of \fig{ck}.)

This pattern of \fig{howxy} (where the changes examples are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
paper. Hence,  we can assert that
predictors learned from these training data have some authority in the regions
used to generate the plans.

@@

That said, the above comes with some important caveats:
\bi
\item 
We   strongly recommend that predictors are assessed prior to planning. That
issue is explored further in the \label{sect:tesd}.
\item
Planners should be designed to be {\em trust increasing}. We list four such planning methods in \tion{planners}.
\item
Where possible, planners should be assessed via some external
oracle that can accurately assess new examples. For an example of that kind of analysis,
see the Validity  in \tion{valid}.
\ei

  \section{Test Data}\label{sect:tesd}
To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
  software system   configuration data from by  Siegmund et al.~\cite{sven12}.
  See github.com/ai-se/XTREE\#data for full access to this data.

@@

\section{Four Planning Methods}\label{sect:planners}
In his textbook on empirical methods, Cohen~\cite{cohen95} advises that any supposedly better
algorithm should be baselined against simpler alternatives. Accordingly,
this section described XTREE (which we call Method4) against three  
alternatives: 
  \bi
  \item  
  Method1, Method2 and Method3 use the   top-down
	bi-clustering method described in \fig{where}.C  which recursively divides the
	data in two  using a dimension that captures the greatest variability in the data. 
	\item
 Method4 uses the decision tree learner of \fig{where}.D. 
  \ei
  Note that all our methods have the two properties proposed in \tion{prelim}:
  \bi
  \item They are {\em local learners}; i.e. different test examples
  will be given plans that are specialized  to their particulars;
  \item They  are {\em trust-increasing} in the manner described in \tion{trust};
  i.e. they change changing examples such that they move {\em closer} to the training data.
  \ei

@@

\subsection{Method4 =  XTREE =\\ Deltas in Decision Tree Branches}

@@

{\em Details:} One potential problem with Methods 1,2 and 3 is the {\em unsupervised} nature of
the algorithm used for the initial clustering. Unsupervised methods like WHERE
execute without knowledge of the target class.
{\em Supervised} methods, on the other hand, are allowed to reflect on more knowledge including
the values in the target class. In theory, this extra knowledge can lead to better data mining.
Accordingly, XTREE uses a supervised method to generate its plans.
Specifically, it  uses the decision tree algorithm of \fig{where}.D to divide the data.
Next, XTREE builds plans from the branches of the decision trees. It does so by asking three questions:
\be
\item
What {\em current} branch does a test case fall in?
\item What {\em desired} branch would the test case want to move to?
\item What are the {\em deltas} between current and desired? 
\ee
The last question can be answered by finding the deltas in branches of the decision tree that lead to {\em desired} from {\em current}. For full algorithmic details, refer to \fig{XTREE_bare}.

@@

\begin{figure}[!b]
\small 
\[
\begin{array}{r}
\mathrm{project}\\
\mathrm{data}
\end{array} 
\left\{\begin{array}{l}\mathit{train}
        \left\{\begin{array}{l}
                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; XTREE)}
              \end{array}\right.
       \\
      ~\\
\mathit{test}  
    \left\{\begin{array}{l@{~}l}
           \mathit{before}& =\mathrm{Performance\; scores \hspace{2pt} in \hspace{2pt}}\mathit{test}\\
           \mathrm{\bf if\;}\mathit{before} & >  \mathit{0}\\
           \mathrm{\bf then} &
           \left\{
            \begin{array}{l}
                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
                \mathrm{{\bf return}\;} R=\frac{\mathit{after}}{\mathit{before}}
            \end{array}
          \right.
   \end{array}\right.
\end{array} \right. 
\]

@@

\subsubsection{Studying Repeated Samples}
Our methods contain some stochastic algorithms; e.g. WHERE's selection of ``what example to explore first'' (see \fig{where}.C) and
  XTREE' occasional use of a random guess when deciding what part of a discretized range to include in the plan
  (see \fig{XTREE_bare}). Hence, we report the $R$ values seen in 40 repeated runs
  (with different random number seeds).
The the value 40 was chosen to be  larger than the 30 samples  required
to satisfy the central limit theorem.

@@

The above table has a  left-hand-side  {\bf Rank} column. This column reports if
the values for each method are statistically different and are more than trivially different. 
In this example, table, CD and BIC are ranked together while XTREE and CD+FS are ranked best and worst, respectively.

@@

XTREE performs well in terms of  effectiveness. 
\fig{jur} and \fig{conf} show the effectiveness of our methods seen in 40 repeats with each data set.
In these experiments,   the dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively. Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data.
\bi
\item For Siegmund data, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data. 
\item For the Jureczko data, we used the training and testing sets of \fig{j}. For these datasets,
 all the SMOTE-ing and Random Forst tunings (discussed in \tion{tesd})
occurred in the {\em train} phase of \fig{work}.
\ei
The results show planning can be very effective-- some of the expected value of the reductions are very large:
\bi
\item
Down to 22\% of the original baseline in Ant of \fig{jur};
\item
Down to 22\% of the original baseline in X264  of \fig{conf};
\ei
In all these results,  with one exception, XTREE always is the top-ranked method and with a median improvement at least 10\% lower (i.e. better) than the next ranked method. The  exception is SQL which (as shown in \fig{conf}), seems to defy all methods for optimization
so XTREE can only reduce the expected values of the runtimes by 1\%.

@@

As predicted in the above discussion on Method 1 (CD), this method always proposes changes  to most features. Method3 (BIC) often proposes changes to many features, which 
would make its plan tedious to implement. 
Method 2 (CD+FS)
proposes less, due to its use of feature selection and, in the case of {\em BDBJ}, it fixates on changing just one feature (which
probably explains why CD+FS performs so poorly for {\em BDBJ} in \fig{conf}). Finally, Method4 (XTREE) rarely fixates
on changing one feature (only on {\em LUCENE}'s lines of code feature) and seems to prefer making many small changes,
all over the space of features.

\fig{changed} reports the changes just for   {\em lucene}) and {\em BDBJ}. \fig{types} shows the mean changes for features found by our
four methods (here, mean change is the average height of the histogram bars seen in, say, \fig{changed}). Overall, Method1 (CD) and
Method3 (BIC) make the most changes while XTREE and CD+FS make the fewest.

\section{Validity}\label{sect:valid}

@@

This main contribution of this paper was to  define an evaluation strategy for this kind of reasoning, and to show that (using that strategy) we
can (a)~select useful planning methods while (b)~culling clearly
inferior planning methods. 
We apply this strategy to XTREE and three alternate planning methods to find:
\bi
\item Two methods that are clearly inferior;
\item One method that proposes complex changes, which work as well as XTREE;
\item And XTREE, that proposes very simple changes.
\ei
Of course, there are many more methods for generating plans and
no   one paper can survey them all.
But the goal of this paper is not to claim that (e.g.) XTREE is some absolute optimal algorithm. Rather, it is
to offer a baseline result (with XTREE) and an  evaluation strategy that  can assess  if alternate methods are better than XTREE.
Our hope is that other   researchers apply this strategy  to repeat and/or improve
and/or refute our results. To that end, all our materials and data sets are available on-line, see
github.com/ai-se/XTREE.

@@

Using DTREE, the test cases can be categorized into one of the branches of the tree. Now, to generate the contrast sets, we determine (1) What \textit{current} branch does a test case fall in?; (2) What \textit{desired} branch would the test case have to move to?; (3) What are the deltas between \textit{current} and \textit{desired}? The last question can be answered by finding the deltas in branches of the decision tree that lead to \textit{desired} from \textit{current}. For full algorithmic details, refer to \fig{XTREE_bare}.

A tree structure such as DTREE is characterized by attributes such as size and depth. It is worth noting that these attributes have a profound impact on its performance. Our initial motivation for using DTREE was that they could serve as a medium for experts to reason about a data. The size of the tree, if too large, jeopardizes the readability of the solutions by increasing the complexity. We have, therefore, endeavored to reduce the size of the tree by pruning away irrelevant solutions that do not contribute to better solutions (refer to step-2 of \fig{XTREE_bare}).
